Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-04-23 paper.tex
Incorporated many of Ib and Oles corrections

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.  The document should begin with an
%% abstract, if appropriate.  If one is given and should not be, the
%% contents will be gobbled.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  High-quality mass spectrometry data of the oscillatory behavior
  of CO oxidation on SiO$_2$ supported Pt-nanoparticles at atmospheric
  pressure has been acquired as a function of pressure, coverage, gas
  composition and nanoparticle size. The oscillations are self-sustained
  for several days at constant temperature, pressure and CO/O$_2$ ratio.
  The frequency of the oscillations is very well defined and increases
  over time. The oscillation frequency is furthermore strongly temperature
  dependent with increasing temperature resulting in increasing frequency.
  A plausible mechanism for the oscillations is proposed based on a
  oxidation/reduction cycle of the nanoparticles which change the rate of
  CO oxidation on the particles.
\end{abstract}

@@

Self-sustained oscillating systems are widespread in nature and have been
described both experimentally and theoretically. One of the maybe most famous
biological examples is the oscillation of the lynx-hare population as reported
through the number furs delivered to the Hudson Bay
Company\cite{MacLulich,Elton1942}. This was for some time believed to be forced
by sunspot oscillations\cite{Elton1924}. Even though the correlation between
sunspots and the lynx-hare population has been debated, even as recently as
1993\cite{SINCLAIR1993}, this is not generally believed to be the case
\cite{Elton1942,Kielland2010}. Instead, the periodic population cycle
can be explained by more classic weather phenomena. Similarly, very interesting
observations have been made for our current climate
changes\cite{Svensmark1997,Svensmark1998} suggesting that the sunspot activity
influences the level of cosmic rays and thus the cloud formation. This topic
is, however, a topic of controversy\cite{Lockwood2007}.

A beautiful example of self-sustained oscillatory behavior in chemistry is the
colorful Belousov- Zhabotinsky reaction\cite{ZAIKIN1970} which is rather easy
to demonstrate for an audience. Such self-sustained oscillatory behavior has
been explained for systems far from equilibrium in several theoretical
works\cite{HakenBog,NicolisBog}. Similar behavior has been observed within
heterogeneous catalysis where one of very first examples was observed by Wicke
for CO oxidation on a catalyst
of highly dispersed Pt on $\gamma$-Al$_2$O$_3$\cite{BEUSCH1972}. To gain a
deeper insight into the fundamental aspects of such oscillatory reactions this
work was followed up by Ertl on Pt(110) surfaces under well-defined conditions
\cite{EISWIRTH1986}. Ertl demonstrated how the switching between the
unreconstructed Pt(110) surface and the reconstructed 2x1-Pt(110) surface
combined with large changes of oxygen sticking coefficient on the two surfaces
could drive the self-sustained oscillations. This and many other interesting
aspects of self-sustained and forced oscillatory behavior and spatio-temporal
self-organization were investigated by Ertl and coworkers on well-defined
single crystals, see e.g. reviews\cite{IMBIHL1995,JAKUBITH1990} and in
particular the Nobel Lecture\cite{Ertl2008}. One of the reasons this has
attracted interest is the importance of such catalytic reactions demanding a
better understanding of their behavior with the goal of further improvement.

CO oxidation is one of the most studied reactions due to its importance in both
automotive catalysis and in the removal of trace CO in hydrogen
streams\cite{IbsBog,Zhang2011}. The oscillatory behavior is not, however, limited to
either Pt or CO oxidation but have been found for many other metallic catalysts
like Rh, Ir, and Pd for oxidation reactions in general, NO and CO reactions,
hydrogenation reactions \cite{Lund2000,SALES1982} and redox
systems\cite{Ibele2010}. Most of the studies
are made on single crystal or polycrystalline films or filaments mainly in
conjunction with ultra-high vacuum equipment to be able to characterize
catalyst surfaces using surface science methods. The amount of work described
on supported catalysts seems more limited due to the difficulty of
characterizing those systems and because additional complexities are present
such as mass and heat transport\cite{Meunier2008}. Reports in the literature
exist of how large
scale facilities such as an industrial ammonia fixed-bed synthesis reactor in
Germany in 1989 started oscillating for several hours. This could be modeled by
coupling of the exothermic process and the heat management in the reactor
system \cite{Morud1998} but one cannot rule out that the observed oscillations
could in some systems also be created by inherent processes on the catalyst
itself. In general such oscillations would be considered highly undesirable in
industry since they would be difficult to control and result in run-away
incidents for, in particular, exothermic reactions. This would result in
damaged catalysts due to sintering at higher temperatures. On the other hand
examples exist in literature where forced oscillations seem to improve the
overall activity\cite{IMBIHL1995,Machado2005}. One particular interesting
process here could be the methanol synthesis where a reduction of the CO$_2$
content for a shorter period of time results in a subsequent higher methanol
rate compensating for a lower methanol production while running in a
reducing H$_2$ and CO mixture alone\cite{Dynamics-Bog}. This enhancement can
be explained by an extensive wetting and un-wetting of the active Cu catalyst
dependent on the oxidation potential\cite{Hansen2002}, i.e. the active area is
strongly increased when reduced but slowly returns to the less active state
when CO$_2$ is present.

The understanding of self-sustained oscillations on supported
catalysts is hampered by the fact that entirely different mechanisms than those
observed on single and polycrystalline samples may exist. For example it was
shown on a Pt(110) crystal how waves of surface reconstruction could propagate
on the surface ensuring a synchronization of the phenomena. Synchronization also have to
occur for oscillations on supported catalysts but here it is more likely to be
due to coupling of mass or heat transport. There have been a few examples where
such oscillations have been observed. Of these the most prominent work is by
Skoghlund\cite{Carlsson2006} where self-sustained oscillations were observed on
a supported Pt catalyst while simultaneously studying the catalyst by in-situ
IR spectrometry. Similar studies were recently undertaken using EXAFS
\cite{Singh2010} but here the oscillatory behavior was unstable and less
convincing.

In this work we demonstrate how a two-dimensional model system of Pt
nanoparticles on a SiO$_2$ support can be constructed that shows highly
repeatable self-sustained oscillations when initially started. The system
consists of mass selected Pt nanoparticles in the range from 2 to 10\,nm 
deposited in a microreactor with a surface area of 0.75\,cm$^2$
especially designed to achieve high sensitivity\cite{Henriksen2009}. This
method offers some advantages over the chemical synthesis methods, which also
offers the possibility of size selected nanoparticles, see e.g.
\cite{Tsang2008,Shao2011}. Particularly, surfactants or solvents
\cite{Biacchi2011}, which are required for
chemical synthesis methods, are avoided with gas-aggregation formed particles
while still being able to vary the coverage from sub-percent to full coverage.
It is demonstrated how the nanoparticles provides self-sustained
oscillations stable in course of several days and how various parameters can be mapped out
including the temperature dependence -- an effect that, to our knowledge, has
never been measured before and that could
not be demonstrated for a continuous film of same area and similar conditions.
By combining these results with previous studies \cite{Carlsson2006,Singh2010}
and a recent detailed study of CO oxidation on Pd using STM
\cite{Hendriksen2010} a plausible explanation for the oscillatory behavior is
presented. The low coverage and extremely efficient Pt nanoparticles illustrate
some of the pit falls that may have to be considered when making in-situ
investigations of such a system since extremely low coverages may provide the
entire conversion.

\section{Results and discussion}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off
CO-oxidation measurement is performed (shown in \ref{fgr:initial_treatment}) where
the temperature is increased at 3\,$^\circ$C/min until the sample ignites and
achieves full conversion of CO to CO$_2$. At a CO:O$_2$ ratio of $\sim0.08$ and a
geometrical Pt coverage of the reactor area of $\sim0.1\%$ light-off occurs at approximately
180\,$^\circ$C. The temperature ramp is typically continued up to
260\,$^\circ$C whereafter the temperature is decreased at the same rate as the
heating ramp. After the initial light-off the sample is cooled to room
temperature. When at room temperature the temperature is again increased at
3\,$^\circ$C/min until sustained oscillations occur. A typical initial
treatment is seen in \ref{fgr:initial_treatment} where both the symmetrical
light-off ramp and the subsequent increase in temperature to achieve sustained
oscillations is seen. Often the sample will show oscillations either directly
upon light-off or on the falling temperature ramp.

\begin{figure}
  \includegraphics[width=12cm]{initial_treatment.png} 
  \caption{A typical example of the initial treatment of a new sample. After
  performing a light- off ramp where oscillations can been seen the temperature
  is increased to a constant value where sustained oscillations take place. The
  non-zero value of CO during high conversion periods is consistent with the
  expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{figure}

The oscillations on all samples are qualitatively similar but the oscillation
period varies from a few seconds (the time constant of the reactor is
$\sim7\,$s) up to more than an hour. Once a sample start to oscillate it will
perform self-sustained oscillations for as long as the experiment is allowed to
run. We have not observed a single sample stop oscillating once the self-sustained oscillations had started.

An example of a single very slow oscillation is seen in
\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where
all CO is converted to CO$_2$ the reaction shows a fast deactivation. 
Notice the CO signal does not go to zero due to the cracking pattern of
CO$_2$ and CO background in the Quadrupole Mass Spectrometer (QMS).
After the deactivation the sample quickly regains
$\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases
slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and convert all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

Due to the long timescales between full conversion and low conversion plateaus
the oscillations cannot be attributed to the reactor setup itself which due to
its small size has timescales on the order of seconds. Furthermore, a series of
experiments were performed to exclude the oscillations as an instrumental
artifact. Firstly, experiments were performed on a larger volume reactor thus
increasing the residence time in the reactor. Secondly, the oscillations were
reproduced on a comparable but not identical setup. Finally, attempts at
reproducing the phenomenon on Pt thin films of comparable coverage did not
produce any oscillations. All of these experiments suggest that the
oscillations is a property of the Pt nanoparticles. However, 
although oscillations have previously been observed on extended Pt thin films
\cite{Singh2010} as well as single crystals\cite{Hendriksen2005} at atmospheric
pressure no oscillations could be provoked on thin films of comparable
geometrical coverage to the nanoparticles in our system at any CO/O$_2$ ratio.

@@

After prolonged measurements over several days an increase in oscillation
period was observed as shown in \ref{fgr:long_measurement}. Initially, the
oscillation period is $\sim$2.5 min and increases linearly with time until 1500\,min
of total oscillation time. Hereafter the oscillations become more irregular
as shown in \ref{fgr:long_measurement} and in the lower panel of \ref{fgr:extracts}. At
experiment end the oscillation period is between 20\,min and 45\,min.

\begin{figure}
  \includegraphics[width=12cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample
  oscillating under constant temperature, pressure and reactant composition.
  The sample went through a total of 439 oscillations in 4 days. Time is
  defined from experiment start and hence includes initial treatment. The inset
  shows the initial steady increase in oscillation period.}
  \label{fgr:long_measurement}
\end{figure}

\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations
extracted from the same four days measurement. Data from shortly after the
initial treatment, after 2800\,min of oscillation time and after 5900\,min is
shown. After 5900\,min the oscillations are more irregular which is consistent
with the data shown in \ref{fgr:long_measurement}. Also, oscillations in
between full conversion oscillations with smaller amplitude and much higher
frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}
  \includegraphics[width=12cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from a 4 day long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.} \label{fgr:extracts}
\end{figure}

The oscillation period increases with time and becomes increasingly less stable
as shown in \ref{fgr:extracts}. This phenomena is  consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more
short-period oscillations despite the trend of increasingly slower oscillation
period. A possible explanation of the phenomeonon could be sintering of the
particles during the oxidation and reductions cycles. However, the ratio
between CO and CO$_2$ integrated over an full period remains constant
throughout the entire experiment time, i.e. the samples maintain a constant
average rate independent of oscillation frequency. Thus we do not expect
sintering to be the cause since an average loss in activity of the sample would
be expected.

It has previously been shown using STM \cite{Hendriksen2002}, FT-IR
\cite{Carlsson2006}, Monte Carlo simulations \cite{Zhdanov2002} as well as DFT
\cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric
pressures is highly dependent on the oxidation state of the surface. A detailed
proposal for the reaction mechanism on Pd has been developed by Hendriksen
\textit{et. al.}\cite{Hendriksen2010} and our data is consistent with this
model.

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal
surface is less active than the oxidized surface. Oscillations originate from a
switch between an oxidized surface and the bare metal surface. Initially, with
CO present in the inlet gas a smooth metal surface is exposed due to the
general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles
convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt
oxide will start to form in the high partial pressures of O$_2$ thus increasing
the rate. The model\cite{Hendriksen2010} suggests that the oxidized
nanoparticles will roughen during CO oxidation resulting in the formation of an
increasingly rough oxide surface. As the oxide becomes rougher the bare metal
surface is increasingly favored resulting in a return to the lower rate rough
metal state. As the rate decreases the CO partial pressure increases which
results in a transition from rough metal surface to a smooth metal surface.
When the sample is sufficiently smooth it will again oxidize and thus complete
the oscillation cycle.

\subsection{Temperature dependence}
\begin{figure}
\includegraphics[width=12cm]{temperature_dependence.png}
\caption{Illustration of the strong temperature dependence. After about 10
  hours of steady oscillations at 205\,$^\circ$C the temperature is lowered by
  5\,$^\circ$C immediately decreasing the oscillation frequency. The temperature
  is hereafter increased by 2\,$^\circ$C resulting in an increase in oscillation
  frequency. At 205\,$^\circ$C the oscillation period was approximately 500\,s
  while the oscillation period period is approximately 5000\,s. at
  203\,$^\circ$C
  \label{fgr:temperature_dependence}
}
\end{figure}

An interesting feature in the data is the very large temperature dependence on
oscillation frequency as shown in \ref{fgr:temperature_dependence}. By
changing the temperature 5\,$^\circ$C the oscillation period was changed by more
than a factor of 10. The magnitude of the change is not completely consistent
across all measured samples but all samples show a very large temperature
dependence where in some cases a provoked small change in temperature will
consistently turn on and off the oscillations, see \ref{fgr:temperature_dependence_supplemental}.
\begin{figure}
\centering
  \includegraphics[width=12cm]{temperature_dependence_supplemental.png}
  \caption{Another example of the very pronounced temperature dependence.
  here the oscillations are turned on and off by changing the temperature
  $20^\circ$C,}
  \label{fgr:temperature_dependence_supplemental}
\end{figure}
This property agrees well an earlier suggested model \cite{Hendriksen2010} and
it is, to our knowledge, the first time the temperature dependence of
oscillations at atmospheric pressure has been measured.

\subsection{Pressure dependence}
The pressure dependence of the oscillations was also investigated. In the
pressure range of 0.1\,bar to 1\,bar no change of the oscillation frequency or
qualitative behavior that could be attributed to the pressure was found.
Similarly, only a weak dependence on the CO/O$_2$ ratio was observed. At a
CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations was
observed. Increasing the CO/O$_2$ showed clear oscillations with a significant
increase in oscillation amplitude compared to lower CO concentrations.
Generally, a tendency towards longer oscillation periods for higher CO/O$_2$
ratio was observed. Richer CO mixtures than 0.175 did not produce any
oscillations in our system.

@@

\subsection{CO concentration dependence}
The dependence of the CO concentration on the period of the oscillations was
investigated. The result shows that the period increases slightly with
increasing CO-concentration. However, the effect is small compared to the
general trend of slower oscillations as the experiment progresses. In
\ref{fgr:gas_dependence_summary} the oscillation periods are summarized and the
actual masspectrometry data of the entire experiment is shown in
\ref{fgr:gas_dependence}.

\begin{figure}
  \includegraphics[width=12cm]{oscillations_gas_dependence_summary_supplemental.png}
  \caption{Summary of the gas-dependence. The overall trend of increasing
  oscillation period is superimposed with small discontinuous steps when the CO
  concentration is increased.} \label{fgr:gas_dependence_summary}
\end{figure}

>>>>

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.  The document should begin with an
%% abstract, if appropriate.  If one is given and should not be, the
%% contents will be gobbled.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  High-quality mass spectrometry data of the oscillatory behavior
  of CO oxidation on SiO$_2$ supported mass selected Pt-nanoparticles at atmospheric
  pressure has been acquired as a function of pressure, coverage, gas
  composition and nanoparticle size. The oscillations are self-sustained
  for several days at constant temperature, pressure and CO/O$_2$ ratio.
  The frequency of the oscillations is very well defined and increases
  over time. The oscillation frequency is furthermore strongly temperature
  dependent with increasing temperature resulting in increasing frequency.
  A plausible mechanism for the oscillations is proposed based on a
  oxidation/reduction cycle of the nanoparticles which change the rate of
  CO oxidation on the particles.
\end{abstract}

@@

Self-sustained oscillating systems are widespread in nature and have been
described both experimentally and theoretically. One of the maybe most famous
biological examples is the oscillation of the lynx-hare population as reported
through the number furs delivered to the Hudson Bay
Company\cite{MacLulich,Elton1942}. This was for some time believed to be forced
by sunspot oscillations\cite{Elton1924}. Even though the correlation between
sunspots and the lynx-hare population has been debated, even as recently as
1993\cite{SINCLAIR1993}, this is not generally believed to be the case
\cite{Elton1942,Kielland2010}. Nevertheless, very interesting oscillation
observations have been made for our current climate
changes\cite{Svensmark1997,Svensmark1998} suggesting that the sunspot activity
influences the level of cosmic rays and thus the cloud formation. This topic
is, however, a topic of controversy\cite{Lockwood2007}.

A beautiful example of self-sustained oscillatory behavior in chemistry is the
colorful Belousov- Zhabotinsky reaction\cite{ZAIKIN1970} which is rather easy
to demonstrate for an audience. Such self-sustained oscillatory behavior has
been explained for systems far from equilibrium in several theoretical
works\cite{HakenBog,NicolisBog}. Similar behavior has been observed within
heterogeneous catalysis where one of the very first examples was observed by Wicke
for CO oxidation on a catalyst
of highly dispersed Pt on $\gamma$-Al$_2$O$_3$\cite{BEUSCH1972}. To gain a
deeper insight into the fundamental aspects of such oscillatory reactions this
work was followed up by Ertl on Pt(110) surfaces under well-defined conditions
\cite{EISWIRTH1986}. Ertl demonstrated how the switching between the
unreconstructed Pt(110) surface and the reconstructed 2x1-Pt(110) surface
combined with large changes of oxygen sticking coefficient on the two surfaces
could drive the self-sustained oscillations. This and many other interesting
aspects of self-sustained and forced oscillatory behavior and spatio-temporal
self-organization were investigated by Ertl and coworkers on well-defined
single crystals, see e.g. reviews\cite{IMBIHL1995,JAKUBITH1990} and in
particular the Nobel Lecture\cite{Ertl2008}. One of the reasons this has
attracted interest is the importance of such catalytic reactions demanding a
better understanding of their behavior with the goal of further improvement.

CO oxidation is one of the most studied reactions due to its importance in both
automotive catalysis and in the removal of trace CO in hydrogen
streams\cite{IbsBog,Zhang2011}. The oscillatory behavior is not, however, limited to
either Pt or CO oxidation but have been found for many other metallic catalysts
like Rh, Ir, and Pd for oxidation reactions in general, NO and CO reactions,
hydrogenation reactions \cite{Lund2000,SALES1982} and redox
systems\cite{Ibele2010}. Most of the studies
are made on single crystal or polycrystalline films or filaments mainly in
conjunction with ultra-high vacuum equipment to be able to characterize
catalyst surfaces using surface science methods. The amount of work described
on supported catalysts seems more limited due to the difficulty of
characterizing those systems and because additional complexities are present
such as mass and heat transport\cite{Meunier2008}. Reports in the literature
exist of how large
scale facilities such as an industrial ammonia fixed-bed synthesis reactor in
Germany in 1989 started oscillating for several hours. This could be modeled by
coupling of the exothermic process and the heat management in the reactor
system \cite{Morud1998} but one cannot rule out that the observed oscillations
could in some systems also be created by inherent processes on the catalyst
itself. In general such oscillations would be considered highly undesirable in
industry since they would be difficult to control and result in run-away
incidents for, in particular, exothermic reactions. This would result in
damaged catalysts due to sintering at higher temperatures. On the other hand
examples exist in literature where forced oscillations seem to improve the
overall activity\cite{IMBIHL1995,Machado2005}. One particular interesting
process here could be the methanol synthesis where a reduction of the CO$_2$
content for a shorter period of time results in a subsequent higher methanol
rate compensating for a lower methanol production while running in a
reducing H$_2$ and CO mixture alone\cite{Dynamics-Bog}. This enhancement can
be explained by an extensive wetting and un-wetting of the active Cu catalyst
dependent on the oxidation potential\cite{Hansen2002}, i.e. the active area is
strongly increased when reduced but slowly returns to the less active state
when CO$_2$ is present. This suggest that an increased activity could be
obtained by oscillating between oxidizing and reducing conditions or CO$_2$
rich/lean conditions respectively.

The understanding of self-sustained oscillations on supported
catalysts is hampered by the fact that entirely different mechanisms than those
observed on single and polycrystalline samples may exist. For example it was
shown on a Pt(110) crystal how waves of surface reconstruction could propagate
on the surface ensuring a synchronization of the phenomena. Synchronization also have to
occur for oscillations on supported catalysts but here it is more likely to be
due to coupling of mass or heat transport. There have been a few examples where
such oscillations have been observed. Of these the most prominent work is by
Skoghlund\cite{Carlsson2006} where self-sustained oscillations were observed on
a supported Pt catalyst while the catalyst was simultaneously studied by in-situ
IR spectrometry. Similar studies were recently undertaken using EXAFS
\cite{Singh2010} but here the oscillatory behavior was unstable and less
convincing.

In this work we demonstrate how a two-dimensional model system of Pt
nanoparticles on a SiO$_2$ support can be constructed that shows highly
repeatable self-sustained oscillations when initially started. The system
consists of mass selected Pt nanoparticles in the range from 2 to 10\,nm 
deposited in a microreactor with a surface area of 0.75\,cm$^2$
especially designed to achieve high sensitivity\cite{Henriksen2009}. This
method offers some advantages over the chemical synthesis methods, which also
offers the possibility of size selected nanoparticles, see e.g.
\cite{Tsang2008,Shao2011}. Particularly, surfactants or solvents
\cite{Biacchi2011}, which are required for
chemical synthesis methods, are avoided with gas-aggregation formed particles
while it is still possible to vary the coverage from sub-percent to full coverage.
It is demonstrated how the nanoparticles provides self-sustained
oscillations stable in course of several days and how various parameters can be mapped out
including the temperature dependence -- an effect that, to our knowledge, has
never been measured before and that could
not be demonstrated for a continuous film of same area in similar conditions.
By combining these results with previous studies \cite{Carlsson2006,Singh2010}
and a recent detailed study of CO oxidation on Pd using STM
\cite{Hendriksen2010} a plausible explanation for the oscillatory behavior is
presented. The low coverage and extremely efficient Pt nanoparticles illustrate
some of the pit falls that may have to be considered when making in-situ
investigations of such a system since extremely low coverages may provide the
entire conversion.

\section{Results and discussion}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off
CO-oxidation measurement is performed (shown in Figure~\ref{fgr:initial_treatment}) where
the temperature is increased at 3\,$^\circ$C/min until the sample ignites and
achieves full conversion of CO to CO$_2$. At a CO:O$_2$ ratio of $\sim0.08$ and a
geometrical Pt coverage of the reactor area of $\sim0.1\%$ light-off occurs at approximately
180\,$^\circ$C. The temperature ramp is typically continued up to
260\,$^\circ$C whereafter the temperature is decreased at the same rate as the
heating ramp. After the initial light-off the sample is cooled to room
temperature. When at room temperature the temperature is again increased at
3\,$^\circ$C/min until sustained oscillations occur. A typical initial
treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical
light-off ramp and the subsequent increase in temperature to achieve sustained
oscillations is seen. Often the sample will show oscillations either directly
upon light-off or on the falling temperature ramp.

\begin{figure}
  \includegraphics[width=12cm]{initial_treatment.png} 
  \caption{A typical example of the initial treatment of a new sample. After
  performing a light-off ramp, where oscillations can been seen, the temperature
  is increased to a constant value where sustained oscillations take place. The
  non-zero value of CO during high conversion periods is consistent with the
  expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{figure}

The oscillations on all samples are qualitatively similar but the oscillation
period varies from a few seconds (the time constant of the reactor is
$\sim7\,$s) up to more than an hour. Once a sample starts to oscillate it will
perform self-sustained oscillations for as long as the experiment is allowed to
run. We have not observed a single sample stop oscillating once the self-sustained oscillations had started.

An example of a single very slow oscillation is seen in
Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where
all CO is converted to CO$_2$ the reaction shows a fast deactivation. 
Notice the CO signal does not go to zero due to the cracking pattern of
CO$_2$ and CO background in the Quadrupole Mass Spectrometer (QMS).
After the deactivation the sample quickly regains
$\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases
slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and converts all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

Due to the long timescales between full conversion and low conversion plateaus
the oscillations cannot be attributed to the reactor setup itself which due to
its small size has timescales on the order of seconds. Furthermore, a series of
experiments were performed to exclude the oscillations as an instrumental
artifact. Firstly, experiments were performed on a reactor with five times larger volume and thus
five times increased residence time in the reactor. Secondly, the oscillations were
reproduced on a comparable but not identical setup. Finally, attempts at
reproducing the phenomenon on Pt thin films of comparable coverage did not
produce any oscillations. All of these experiments suggest that the
oscillations is a property of the Pt nanoparticles. However, 
although oscillations have previously been observed on extended Pt thin films
\cite{Singh2010} as well as single crystals\cite{Hendriksen2005} at atmospheric
pressure no oscillations could be provoked on thin films of comparable
geometrical coverage to the nanoparticles in our system at any CO/O$_2$ ratio.

@@

After prolonged measurements over several days an increase in oscillation
period was observed as shown in Figure~\ref{fgr:long_measurement}. Initially, the
oscillation period is $\sim$2.5 min and it increases linearly with time until 1500\,min
of total oscillation time. Hereafter the oscillations become more irregular
as shown in Figure~\ref{fgr:long_measurement} and in the lower panel of Figure~\ref{fgr:extracts}. At
experiment end the oscillation period is between 20\,min and 45\,min.

\begin{figure}
  \includegraphics[width=12cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample
  oscillating under constant temperature, pressure and reactant composition.
  The sample went through a total of 439 oscillations in 4 days. For details on on
  the oscillations over this period please consult supplementary Figure ????. Time is
  defined from experiment start and hence includes initial treatment. The inset
  shows the initial steady increase in oscillation period. The red lines indicate
  max and min oscillation time to guide the eye.}
  \label{fgr:long_measurement}
\end{figure}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations
extracted from the same four days measurement. Data from shortly after the
initial treatment, after 2800\,min of oscillation time and after 5900\,min are
shown. After 5900\,min the oscillations are more irregular which is consistent
with the data shown in Figure~\ref{fgr:long_measurement}. Also, oscillations in
between full conversion oscillations with smaller amplitude and much higher
frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}
  \includegraphics[width=12cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from a 4 days long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.} \label{fgr:extracts}
\end{figure}

The oscillation period increases with time and becomes increasingly less stable
as shown in Figure~\ref{fgr:extracts}. This phenomenon is consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more
short-period oscillations despite the trend of increasingly longer oscillation
period. A possible explanation of the phenomenon could be sintering of the
particles during the oxidation and reductions cycles. However, the ratio
between CO and CO$_2$ integrated over an full period remains constant
throughout the entire experiment time, i.e. the samples maintain a constant
average rate independent of oscillation frequency. Thus we do not expect
sintering to be the cause since an average loss in activity of the sample would
be expected.

It has previously been shown using STM \cite{Hendriksen2002}, FT-IR
\cite{Carlsson2006}, Monte Carlo simulations \cite{Zhdanov2002} as well as DFT
\cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric
pressures is highly dependent on the oxidation state of the surface. A detailed
proposal for the reaction mechanism on Pd has been developed by Hendriksen
\textit{et. al.}\cite{Hendriksen2010} and our data are consistent with this
model.

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal
surface is less active than the oxidized surface. Oscillations originate from a
switch between an oxidized surface and the bare metal surface. Initially, with
CO present in the inlet gas a smooth metal surface is exposed due to the
general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles
convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt
oxide will start to form in the high partial pressure of O$_2$ thus increasing
the rate. The model\cite{Hendriksen2010} suggests that the oxidized
nanoparticles will roughen during CO oxidation resulting in the formation of an
increasingly rough oxide surface. As the oxide becomes rougher the bare metal
surface is increasingly favored resulting in a return to the lower rate rough
metal state. As the rate decreases the CO partial pressure increases which
results in a transition from rough metal surface to a smooth metal surface.
When the sample is sufficiently smooth it will again oxidize and thus complete
the oscillation cycle.

\subsection{Temperature dependence}
\begin{figure}
\includegraphics[width=12cm]{temperature_dependence.png}
\caption{Illustration of the strong temperature dependence of the oscillation period. After about 10
  hours of steady oscillations at 205\,$^\circ$C the temperature is lowered by
  5\,$^\circ$C immediately decreasing the oscillation frequency. The temperature
  is hereafter increased by 2\,$^\circ$C resulting in an increase in oscillation
  frequency. At 205\,$^\circ$C the oscillation period was approximately 500\,s
  while the oscillation period period is approximately 5000\,s. at
  203\,$^\circ$C.
  \label{fgr:temperature_dependence}
}
\end{figure}

An interesting feature in the data is the very large temperature dependence of the
oscillation frequency as shown in Figure~\ref{fgr:temperature_dependence}. By
changing the temperature 5\,$^\circ$C the oscillation period was changed by more
than a factor of 10. The magnitude of the change is not completely consistent
across all measured samples but all samples show a very large temperature
dependence. In some cases a provoked change in temperature will
consistently turn on and off the oscillations. This is illustrated in the supplementary
material, Figure ??????\ref{fgr:temperature_dependence_supplemental}.
\begin{figure}
\centering
  \includegraphics[width=12cm]{temperature_dependence_supplemental.png}
  \caption{Another example of the very pronounced temperature dependence of the oscillation period.
  Here the oscillations are turned on and off by changing the temperature by
  $20^\circ$C.}
  \label{fgr:temperature_dependence_supplemental}
\end{figure}
This property agrees well an earlier suggested model \cite{Hendriksen2010} and
it is, to our knowledge, the first time the temperature dependence of
oscillations at atmospheric pressure has been measured.

\subsection{Pressure dependence}
The pressure dependence of the oscillations was also investigated. In the
pressure range of 0.1\,bar to 1\,bar no change of the oscillation frequency or
qualitative behavior that could be attributed to the pressure was found.
Similarly, only a weak dependence on the CO/O$_2$ ratio was observed. At a
CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations were
observed. Increasing the CO/O$_2$ ratio showed clear oscillations with a significant
increase in oscillation amplitude compared to lower CO concentrations.
Generally, a tendency towards longer oscillation periods for higher CO/O$_2$
ratio was observed. Richer CO mixtures than 0.175 did not produce any
oscillations in our system.

@@

\subsection{CO concentration dependence}
The dependence of period of the oscillations on the the CO concentration was
investigated. The result shows that the period increases slightly with
increasing CO-concentration. However, the effect is small compared to the
general trend of slower oscillations as the experiment progresses. In
Figure~\ref{fgr:gas_dependence_summary} the oscillation periods are summarized and the
actual mass spectrometry data of the entire experiment are shown in Figure~\ref{fgr:gas_dependence}.

\begin{figure}
  \includegraphics[width=12cm]{oscillations_gas_dependence_summary_supplemental.png}
  \caption{Oscillation period (blue) as a function of time and CO concentration (black). The
  overall trend of increasing oscillation period with time is superimposed with small
  discontinuous steps when the CO concentration is increased.} 
  \label{fgr:gas_dependence_summary}
\end{figure}
